---
title: "10_Figures RSS"
author: "Becky"
date: "3/3/2020"
output: html_document
---

Load files needed for creating figures. Leslie filtered the predictions files by Chinook and coho and life stages: rearing, spawning, and present in script 5.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

Predictions across different habitats.

```{r}
#Load predictions data frame.
# load(file = "output/preds.Rdata")

#load chinook rearing and spawning datasets
load(file = "output/K_r.Rdata")
load(file = "output/K_s.Rdata")
load(file = "output/K_p.Rdata")

load(file = "output/CO_r.Rdata")
load(file = "output/CO_s.Rdata")
load(file = "output/CO_p.Rdata")

```

# Empirical data

Create a new variable with bins by watershed size.

```{r}
temp %>% 
  distinct(Site, cont_area) %>% 
  count(cont_area) 

temp %>% 
  mutate(cont_area_fc = case_when())

```


Plots showing range of empirical data, distribution across sites and years.

```{r}
load(file = "output/temp.Rdata")



temp %>% 
  distinct(Site, year)


all.days <- temp %>% complete(day, nesting(Site, cont_area))

temp %>% 
  mutate(site.year = paste(Site, year, sep = ""),
         cont_area_fc = cut(cont_area, seq(0, 600, 150),
                            labels = c("0-150", "151-300", "301-450", "451-579"))) %>% 
  complete(day, nesting(site.year, cont_area_fc)) %>% 
  # filter(site.year == "APU82015") %>%
  select(site.year, meanT, day, cont_area_fc) %>% 
  ggplot(aes(x = day, y = meanT, group = site.year)) +
  geom_line(show.legend = FALSE) +
  facet_wrap(~cont_area_fc) +
  scale_x_continuous(breaks = c(166, 196, 227, 258), 
                     labels = c("June", "July", "August", "September")) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        text = element_text(size = 18)) +
  labs(y = "Mean Daily Temperature (˚C)",
       title = bquote("Stream Temperatures by Watershed Area"~(km^2)))

ggsave("output/Empirical Stream Temp Data by wtd area.jpg", width = 10, height = 8)





```



```{r}

temp %>% 
  count(Site, cont_area, year)

temp %>% 
  mutate(cont_area_fc = cut(cont_area, breaks = seq(0, 600, 150), 
                            labels = c("0-150", "151-300", "301-450", "451-579"))) %>% 
  count(cont_area_fc, year) %>% 
  ggplot(aes(x = cont_area_fc, y = year, color = n)) +
  geom_point(shape = 15, size = 10) +
  theme_bw() +
  theme(axis.title.y = element_blank(), legend.position = "bottom", 
        text = element_text(size = 18)) +
  labs(x = bquote("Contributing area"~(km^2)), color = "Days per year")

ggsave("output/Data Availability.jpg", width = 6, height = 6)

```


# Predictions

Plot of observed v fitted values.

```{r}
load(file = "output/temp.Rdata")

temp %>% 
  left_join(preds %>% select(rca_id, date, preds), by = c("rca_id" = "rca_id", "sampleDate" = "date")) %>% 
  mutate(cont_area_fc = cut(cont_area, breaks = seq(0, 600, 150), 
                            labels = c("0-150", "151-300", "301-450", "451-579")),
         season = case_when(day < 202 ~ "spring",
                            TRUE ~ "fall")) %>% 
  ggplot() +
  geom_point(aes(x = preds, y = meanT, color = season), shape = 1) +
  facet_wrap(~cont_area_fc) +
  theme_bw()
```

Number of sites by watershed size.

```{r}
temp %>% 
  mutate(cont_area_fc = cut(cont_area, breaks = seq(0, 600, 150), 
                            labels = c("0-150", "151-300", "301-450", "451-579"))) %>% 
  distinct(cont_area_fc, Site) %>% 
  count(cont_area_fc)

temp %>% distinct(cont_area, Site) %>% 
  ggplot() +
  geom_point(aes(x = cont_area, y = fct_reorder(Site, cont_area)))
```


Plots that show distributions of temperatures by decade. To show the distributions of temperatures experienced by each species and life stage, need a mean +/- 2 SD. 

```{r}


#Chinook rearing
decade_K_r <- K_r %>% 
  group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) %>% 
  mutate(decade = case_when(year< 1990 ~ 1980, 
                            year >= 1990 & year < 2000 ~ 1990, 
                            year >=2000 & year < 2010 ~ 2000, 
                            year>2009 ~ 2010))

decade_K_r %>%
  ggplot() +
  geom_line(aes(x = day, y = mean_yr, color = as.factor(year)), show.legend = FALSE) +
  # geom_ribbon(aes(x = day, ymin = mean_yr - 1.96 * sd_yr, 
  #                 ymax = mean_yr + 1.96 * sd_yr, fill = as.factor(year)),
  #             alpha = 0.1, show.legend = FALSE) + 
  facet_wrap(~decade) + 
  theme_bw() +
  ggtitle("Mean daily temperatures by year for Chinook rearing habitat") 

ggsave("salmon_figs/Mean daily chinook rearing thermal curve.jpeg", plot=last_plot())


```

Boxplots by day. 

```{r}

K_p %>% 
  filter(day %in% 150:210) %>% 
  ggplot(aes(x = as.factor(day), y = preds)) +
  geom_boxplot() + 
  facet_wrap(~decade)

```

Current baseline (2000-2018) temperature for Chinook rearing.

```{r}



```


# Future scenarios


```{r}
load("output/preds_S1.Rdata")
load("output/preds_S2.Rdata")
load("output/preds_S3.Rdata")
load("output/preds_S4.Rdata")
```


```{r}
K_rcas <- preds %>% 
  filter(K_p == 1) %>% 
  distinct(rca_id, reach_length)


K_rcas

preds_S4 %>% 
  right_join(K_rcas) %>%
  group_by(rca_id, month) %>% 
  summarize(mean_temp = mean())
  ggplot() +
  geom_histogram(aes(x = preds_bl)) 
  geom_histogram(aes(x = preds_fut), color = "red", alpha = 0.1)

```

