---
title: "11_Future Scenarios"
author: "Becky"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(zoo)
library(lubridate)
library(tidyverse)
library(sf)

load("output/preds.Rdata")
load("output/spring.lm.Rdata")
load("output/fall.lm.Rdata")

```


# Future Scenarios

Create a current baseline using predictions for 2000-2018. Baseline includes average values across all RCAs for 3-day air temperatures and 5-day precipitation. Then grab a low snow and high snow year within that baseline. 

New data for predicting stream tempatures:

* Scenario 1: low snow and 2 deg air temperature change.
* Scenario 2: low snow and 4 deg air temperature change.
* Scenario 3: high snow and 2 deg air temperature change.
* Scenario 4: high snow and 4 deg air temperature change.

Identify low and high snow years within the baseline period. Make them similar to the low and high snow years in the historic scenarios. < 10 mm and > 100 mm.

2nd lowest snow year was 2005 (12 mm) and 4th highest snow year was 2011 (113 mm).

```{r}
swe_bl <- preds %>% 
  filter(year %in% 2000:2018) %>% 
  group_by(year) %>% 
  summarize(sweA1 = mean(sweA1))

swe_bl %>% 
  ggplot(aes(y = fct_reorder(as.factor(year), sweA1), x = sweA1)) +
  geom_point() 

swe_bl %>% 
  ggplot() +
  geom_histogram(aes(x = sweA1), bins = 10)

swe_bl %>% 
  arrange(desc(sweA1))

```


Create data frames for prediction under the four scenarios.

```{r}

newData_S1 <- preds %>% 
  filter(year %in% 2000:2018) %>% 
  group_by(rca_id, day, cont_area, reach_slope, rca_elev_mn) %>% 
  summarize(tair3 = mean(tair3) + 2,
            prcp5 = mean(prcp5),
            preds_bl = mean(preds)) %>% 
  left_join(preds %>% filter(year == 2005) %>% distinct(rca_id, sweA1))

newData_S2 <- preds %>% 
  filter(year %in% 2000:2018) %>% 
  group_by(rca_id, day, cont_area, reach_slope, rca_elev_mn) %>% 
  summarize(tair3 = mean(tair3) + 4,
            prcp5 = mean(prcp5),
            preds_bl = mean(preds)) %>% 
  left_join(preds %>% filter(year == 2005) %>% distinct(rca_id, sweA1))

newData_S3 <- preds %>% 
  filter(year %in% 2000:2018) %>% 
  group_by(rca_id, day, cont_area, reach_slope, rca_elev_mn) %>% 
  summarize(tair3 = mean(tair3) + 2,
            prcp5 = mean(prcp5),
            preds_bl = mean(preds)) %>% 
  left_join(preds %>% filter(year == 2011) %>% distinct(rca_id, sweA1))

newData_S4 <- preds %>% 
  filter(year %in% 2000:2018) %>% 
  group_by(rca_id, day, cont_area, reach_slope, rca_elev_mn) %>% 
  summarize(tair3 = mean(tair3) + 4,
            prcp5 = mean(prcp5),
            preds_bl = mean(preds)) %>% 
  left_join(preds %>% filter(year == 2011) %>% distinct(rca_id, sweA1))

```

Predictions for four scenarios.

Function to predict by spring and fall models and bind data together.

```{r}

scenario_predictions <- function(newData_byScenario) {
  spring_dat <- newData_byScenario %>% filter(day < 202)
  spring_preds <- spring_dat %>% 
    ungroup() %>% 
    mutate(preds_fut = predict(spring.lm, newdata = spring_dat))
  fall_dat <- newData_byScenario %>% filter(day > 201)
  fall_preds <- fall_dat %>% 
    ungroup() %>% 
    mutate(preds_fut = predict(fall.lm, newdata = fall_dat))
  preds <- bind_rows(spring_preds, fall_preds) %>% 
    mutate(change = preds_fut - preds_bl,
           date = as.Date(day, origin = "1980-01-01"),
           month = month(date))
  return(preds)
}

```

Predict across four scenarios.

```{r}
preds_S1 <- scenario_predictions(newData_S1)
save(preds_S1, file = "output/preds_S1.Rdata")

preds_S2 <- scenario_predictions(newData_S2)
save(preds_S2, file = "output/preds_S2.Rdata")

preds_S3 <- scenario_predictions(newData_S3)
save(preds_S3, file = "output/preds_S3.Rdata")

preds_S4 <- scenario_predictions(newData_S4)
save(preds_S4, file = "output/preds_S4.Rdata")

```


# NOT DONE - replaced with simpler code to a table.
For each scenario, calculate the mean monthly temperature and the mean monthly temperature change. Include the baseline mean monthly temperature as well for putting in ArcGIS.

Read in reaches feature class from geodb.

```{r}
fgdb <- "T:\\Aquatic\\KFHP\\Geodatabases\\Anchor.gdb"

#merge with spatial data and export
rca_reach <- sf::st_read(dsn = fgdb, layer = "anch_rca_reaches_attributed_06022020") %>%
  mutate(rca_id = reach_id) %>% 
  select(rca_id, Shape, Shape_Length)
```

Create monthly metrics. 

```{r}

baseline_monthly_temps <- preds_S1 %>% 
  filter(month %in% 6:9) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_bl = mean(preds_bl)) %>% 
  spread(key = month, value = mean_bl) %>% 
  rename(bl_6 = "6", bl_7 = "7", bl_8 = "8", bl_9 = "9")

S1_monthly_temps <- preds_S1 %>% 
  filter(month %in% 6:9) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_fut = mean(preds_fut)) %>% 
  spread(key = month, value = mean_fut) %>% 
  rename(S1_6 = "6", S1_7 = "7", S1_8 = "8", S1_9 = "9")

S2_monthly_temps <- preds_S2 %>% 
  filter(month %in% 6:9) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_fut = mean(preds_fut)) %>% 
  spread(key = month, value = mean_fut) %>% 
  rename(S2_6 = "6", S2_7 = "7", S2_8 = "8", S2_9 = "9")

S3_monthly_temps <- preds_S3 %>% 
  filter(month %in% 6:9) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_fut = mean(preds_fut)) %>% 
  spread(key = month, value = mean_fut) %>% 
  rename(S3_6 = "6", S3_7 = "7", S3_8 = "8", S3_9 = "9")

S4_monthly_temps <- preds_S4 %>% 
  filter(month %in% 6:9) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_fut = mean(preds_fut)) %>% 
  spread(key = month, value = mean_fut) %>% 
  rename(S4_6 = "6", S4_7 = "7", S4_8 = "8", S4_9 = "9")

```

Save to shapefiles.

```{r}


rcas_baseline <- sp::merge (rca_reach, baseline_monthly_temps, by="rca_id") 
st_write(rcas_baseline, "output/Shapefiles/rcas_baseline.shp", driver="ESRI Shapefile")

rcas_S1 <- sp::merge (rca_reach, S1_monthly_temps, by="rca_id") 
st_write(rcas_S1, "output/Shapefiles/rcas_S1.shp", driver="ESRI Shapefile")

rcas_S2 <- sp::merge (rca_reach, S2_monthly_temps, by="rca_id") 
st_write(rcas_S2, "output/Shapefiles/rcas_S2.shp", driver="ESRI Shapefile")

rcas_S3 <- sp::merge (rca_reach, S3_monthly_temps, by="rca_id") 
st_write(rcas_S3, "output/Shapefiles/rcas_S3.shp", driver="ESRI Shapefile")

rcas_S4 <- sp::merge (rca_reach, S4_monthly_temps, by="rca_id") 
st_write(rcas_S4, "output/Shapefiles/rcas_S4.shp", driver="ESRI Shapefile")


```


Differences for four scenarios.

```{r}
S1_monthly_diffs <- preds_S1 %>% 
  filter(month %in% 6:9) %>% 
  mutate(diff = preds_fut - preds_bl) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_diff = mean(diff)) %>% 
  spread(key = month, value = mean_diff) %>% 
  rename(S1_6 = "6", S1_7 = "7", S1_8 = "8", S1_9 = "9")

S2_monthly_diffs <- preds_S2 %>% 
  filter(month %in% 6:9) %>% 
  mutate(diff = preds_fut - preds_bl) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_diff = mean(diff)) %>% 
  spread(key = month, value = mean_diff) %>% 
  rename(S2_6 = "6", S2_7 = "7", S2_8 = "8", S2_9 = "9")

S3_monthly_diffs <- preds_S3 %>% 
  filter(month %in% 6:9) %>% 
  mutate(diff = preds_fut - preds_bl) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_diff = mean(diff)) %>% 
  spread(key = month, value = mean_diff) %>% 
  rename(S3_6 = "6", S3_7 = "7", S3_8 = "8", S3_9 = "9")

S4_monthly_diffs <- preds_S4 %>% 
  filter(month %in% 6:9) %>% 
  mutate(diff = preds_fut - preds_bl) %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_diff = mean(diff)) %>% 
  spread(key = month, value = mean_diff) %>% 
  rename(S4_6 = "6", S4_7 = "7", S4_8 = "8", S4_9 = "9")


```

Save to shapefiles.

```{r}

rcas_S1_diff <- sp::merge (rca_reach, S1_monthly_diffs, by="rca_id") 
st_write(rcas_S1_diff, "output/Shapefiles/rcas_S1_diff.shp", driver="ESRI Shapefile")

rcas_S2_diff <- sp::merge (rca_reach, S2_monthly_diffs, by="rca_id") 
st_write(rcas_S2_diff, "output/Shapefiles/rcas_S2_diff.shp", driver="ESRI Shapefile")

rcas_S3_diff <- sp::merge (rca_reach, S3_monthly_diffs, by="rca_id") 
st_write(rcas_S3_diff, "output/Shapefiles/rcas_S3_diff.shp", driver="ESRI Shapefile")

rcas_S4_diff <- sp::merge (rca_reach, S4_monthly_diffs, by="rca_id") 
st_write(rcas_S4_diff, "output/Shapefiles/rcas_S4_diff.shp", driver="ESRI Shapefile")


```

# Table of mean monthly temeperatures for baseline and future scenarios.

Want mean July temperatures for baseline and future scenarios for mapping.

```{r}

mean_monthly_temps_baseline <- preds_S1 %>% 
  group_by(rca_id, month) %>% 
  summarize(baseline = mean(preds_bl)) %>% 
  pivot_wider(names_from = month, values_from = baseline, names_prefix = "baseline_")

mean_monthly_temps_S1 <- preds_S1 %>% 
  group_by(rca_id, month) %>% 
  summarize(S1 = mean(preds_fut)) %>% 
  pivot_wider(names_from = month, values_from = S1, names_prefix = "low2_")

mean_monthly_temps_S2 <- preds_S2 %>% 
  group_by(rca_id, month) %>% 
  summarize(S2 = mean(preds_fut)) %>% 
  pivot_wider(names_from = month, values_from = S2, names_prefix = "low4_")

mean_monthly_temps_S3 <- preds_S3 %>% 
  group_by(rca_id, month) %>% 
  summarize(S3 = mean(preds_fut)) %>% 
  pivot_wider(names_from = month, values_from = S3, names_prefix = "high2_")

mean_monthly_temps_S4 <- preds_S4 %>% 
  group_by(rca_id, month) %>% 
  summarize(S4 = mean(preds_fut)) %>% 
  pivot_wider(names_from = month, values_from = S4, names_prefix = "high4_")


future_July_temps <- left_join(mean_monthly_temps_baseline %>% select(rca_id, baseline_7),
          mean_monthly_temps_S1 %>% select(rca_id, low2_7))
future_July_temps <- left_join(future_July_temps,
          mean_monthly_temps_S2 %>% select(rca_id, low4_7))
future_July_temps <- left_join(future_July_temps,
          mean_monthly_temps_S3 %>% select(rca_id, high2_7))
future_July_temps <- left_join(future_July_temps,
          mean_monthly_temps_S4 %>% select(rca_id, high4_7))

future_July_temps

write.csv(future_July_temps, file = "output/Future_July_temps.csv")

```

Combine all scenarios so can calculate metrics for each.

```{r}
fut_preds <- bind_rows(preds_S1 %>% mutate(scenario = "baseline") %>% 
                         select(rca_id, month, day, scenario, preds = preds_bl),
                       preds_S1 %>% mutate(scenario = "S1") %>% 
                         select(rca_id, month, day, scenario, preds = preds_fut),
                       preds_S2 %>% mutate(scenario = "S2") %>% 
                         select(rca_id, month, day, scenario, preds = preds_fut),
                       preds_S3 %>% mutate(scenario = "S3") %>% 
                         select(rca_id, month, day, scenario, preds = preds_fut),
                       preds_S4 %>% mutate(scenario = "S4") %>% 
                         select(rca_id, month, day, scenario, preds = preds_fut))


fut_mets <- fut_preds %>% 
  group_by(rca_id, scenario) %>% 
  mutate(MA_mean = rollmean(preds, k = 7, align = "center", fill = NA)) %>%  
  summarize(MxDAT = max(preds),
            MxDAT_jd = day[max(preds) == preds][1],
            MA7d_DAT = max(MA_mean, na.rm = TRUE),
            MA7d_DAT_jd = day[max(MA_mean, na.rm = TRUE) == MA_mean][1],
            CDD = sum(preds),
            RNG = max(preds) - min(preds),
            SIGMA_MN = var(preds),
            SUM_13_jas = sum(preds[month %in% 7:9] > 13),
            SIGMA_MN_jas = var(preds[month %in% 7:9]),
            June_mn = mean(preds[month == 6]),
            July_mn = mean(preds[month == 7]),
            Aug_mn = mean(preds[month == 8]),
            Sep_mn = mean(preds[month == 9]))

#check

fut_preds %>% 
  filter(rca_id == 1, scenario == "S1") %>% 
  group_by(rca_id, scenario) %>% 
  mutate(MA_mean = rollmean(preds, k = 7, align = "center", fill = NA))

fut_mets %>% 
  filter(rca_id == 1, scenario == "S1")

```


Save mean July temperature and MA7d_MAT for mapping.

```{r}

left_join(fut_mets %>% 
            select(rca_id, scenario, MA7d_DAT) %>% 
            pivot_wider(names_from = scenario, values_from = MA7d_DAT, names_prefix = "MA7d_DAT_"),
          fut_mets %>% 
            select(rca_id, scenario, July_mn) %>% 
            pivot_wider(names_from = scenario, values_from = July_mn, names_prefix = "July_mn_")) %>% 
  write.csv(file = "output/Future_metrics.csv")

```

Plot of future metrics.

```{r}

fut_mets %>% 
  ggplot() +
  geom_histogram(aes(x = July_mn)) +
  facet_wrap(~scenario)


fut_mets %>% 
  ggplot() +
  geom_density(aes(x = July_mn, color = scenario)) 
```


# Distributions of temperature changes. 

```{r}

scen_change <- bind_rows(preds_S1 %>%
            group_by(rca_id, month) %>%
            summarize(ave_change = mean(change)) %>%
            mutate(scenario = "S1"),
          preds_S2 %>% 
            group_by(rca_id, month) %>% 
            summarize(ave_change = mean(change)) %>%
            mutate(scenario = "S2"),
          preds_S3 %>% 
            group_by(rca_id, month) %>% 
            summarize(ave_change = mean(change)) %>%
            mutate(scenario = "S3"),
          preds_S4 %>% 
            group_by(rca_id, month) %>% 
            summarize(ave_change = mean(change)) %>%
            mutate(scenario = "S4"))

```

Figures of changes.

```{r}

scen_change %>%
  filter(month %in% 6:9) %>% 
  mutate(scenario = factor(scenario, labels = c("Low snow + 2 deg change",
                                                "Low snow + 4 deg change",
                                                "High snow + 2 deg change",
                                                "High snow + 4 deg change"))) %>% 
  ggplot() +
  geom_density(aes(x = ave_change, fill = as.factor(month)), alpha = 0.4) +
  facet_wrap(~scenario) +
  theme_bw() +
  labs(x = "Stream temperature change",
       title = "Changes in stream temperature for four scenarios",
       subtitle = "Average monthly change for all rcas",
       fill = "month")


scen_change %>%
  filter(month %in% 6:9) %>% 
  mutate(scenario = factor(scenario, labels = c("Low snow + 2 deg change",
                                                "Low snow + 4 deg change",
                                                "High snow + 2 deg change",
                                                "High snow + 4 deg change"))) %>% 
  ggplot() +
  geom_density(aes(x = ave_change, fill = as.factor(month)), alpha = 0.4) +
  facet_wrap(~scenario) +
  theme_bw() +
  labs(x = "Stream temperature change",
       title = "Changes in stream temperature for four scenarios",
       subtitle = "Average monthly change for all rcas",
       fill = "month")

ggsave("output/Density plots of changes by scenario.jpg")
 

```

Compare density plots of baseline to future by month for each scenario, using temperatures. Density plots show how distributions are changing. For Chinook rearing, spawning, present.


```{r}

preds_S4 %>% 
  group_by(rca_id, month) %>% 
  summarize(mean_bl = mean(preds_bl),
            mean_fut = mean(preds_fut)) %>% 
  ggplot() +
  geom_density(aes(x = preds_bl, fill = as.factor(month)))

```

