---
title: "2_model exploration"
author: "Leslie Jones"
date: "1/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)

library(lubridate)
library(rgdal)
library(zoo)
library(sf)
library(corrplot)
library(MuMIn)
library(ape)
library(readxl)
#library(VIF)
library(gridExtra)
library(tidyverse)
#library(jtools)
```

# Database for Anchor Model
Load data frame from 1_Create model data frame.Rmd

```{r}
load(file = "output/temp.Rdata")

```


## Compare LST versus daymet air temperatures

See correlation between lst and daymet.
Test models of stream temp with each.

```{r}

cor(temp$lst, temp$tair3, use = "pairwise.complete.obs")

temp %>% 
  ggplot() +
  geom_point(aes(x = tair3, y = lst)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)


```

Look at correlations between each air temperature covariate with stream temperature and also predictors of stream temperature.

```{r}
temp %>% 
  filter(!is.na(lst)) %>% 
  summarize(clst = cor(lst, meanT),
            cdm3 = cor(tair3, meanT),
            cdm8 = cor(tair, meanT))


lm1 <- lm(meanT ~ tair3, data = temp %>% filter(!is.na(lst)))
lm2 <- lm(meanT ~ tair, data = temp %>% filter(!is.na(lst)))
lm3 <- lm(meanT ~ lst, data = temp %>% filter(!is.na(lst)))
AIC(lm1, lm2, lm3)

summary(lm1)
summary(lm2)
```


## Compare 3-day and 8-day daymet air temperatures

This comparison is slightly different from the one above because now we are using all daily stream temperatures. For the LST comparison, we could only use every 8 days to make the data comparable.

```{r}
temp %>% 
  summarize(cdm8 = cor(tair, meanT),
            cdm3 = cor(tair3, meanT))


lm1 <- lm(meanT ~ tair3, data = temp)
lm2 <- lm(meanT ~ tair, data = temp)
AIC(lm1, lm2)

summary(lm1)
summary(lm2)


g1 <- ggplot(data = temp, aes(x = tair3, y = meanT)) +
  geom_point() +
  geom_smooth() +
  geom_abline(aes(slope = 1, intercept = 0))

g2 <- ggplot(data = temp, aes(x = tair, y = meanT)) +
  geom_point() +
  geom_smooth() +
  geom_abline(aes(slope = 1, intercept = 0))

grid.arrange(g1,g2)

```

## Explore correlations between the covariates

Spatial covariates first.


```{r}

temp %>% 
  distinct(reach_slope, reach_length, Forest, Shrub, Wetland, Riparian, 
           rca_elev_min, rca_elev_max, rca_elev_mn, 
           ca_elev_mn, cont_area, elev_delta, sweA1) %>% 
  cor(.) %>% 
  corrplot.mixed(.)

```




## Exploratory Graphs

### Correlation between LST and air temperature by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(lst2, tair, use = "pairwise.complete.obs"))
```

### correlation between stream temp and air temperature by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(ma_mean, tair ))
```

### correlation and plot between stream temp and LST by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(ma_mean, lst2, use = "pairwise.complete.obs"))

temp.dm %>% 
  ggplot() +
  geom_point(aes(x = lst2, y = ma_mean, color = as.factor(month))) +
  facet_wrap(~Site) 
```

### plot air temperature and LST by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>% 
  ggplot() +
  geom_point(aes(x = tair, y = lst2)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)

```
### plot air temperature and stream temperature by month

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
temp.dm %>% 
  ggplot() +
  geom_point(aes(x = tair, y = ma_mean)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)

month5 <- filter(temp.dm,month==5)   
  ggplot() +
  geom_point(data=month5, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
  month6 <- filter(temp.dm,month==6)   
  ggplot() +
  geom_point(data=month6, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
  month7 <- filter(temp.dm,month==7)   
  ggplot() +
  geom_point(data=month7, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
   month8 <- filter(temp.dm,month==8)   
  ggplot() +
  geom_point(data=month8, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
  month9 <- filter(temp.dm,month==9)   
  ggplot() +
  geom_point(data=month9, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")

```


## Exploratory Modeling

Try out sets of models.

```{r, echo=FALSE}

corrplot.mixed(cor(temp %>% select(ma_mean, tair, prcp, swe, slope_P, elev_mean, cont_area, wetland, forest, shrub)))

#model1 <- lm(ma_mean ~ tair, data=temp.mod)
#model2 <- lm(ma_mean ~ tair + elev_mean, data=temp.mod)
#model3 <- lm(ma_mean ~ tair + elev_mean + slope_P, data=temp.mod)
#model4 <- lm(ma_mean ~ tair + elev_mean + slope_P + cont_area, data=temp.mod)
#model5 <- lm(ma_mean ~ tair + elev_mean + slope_P + cont_area + prcp, data=temp.mod)
#AIC(model1, model2, model3, model4, model5)

#make month a factor
month.f <- factor(temp$month)
is.factor(temp$month.f)

full.model <- lm(ma_mean ~ tair + month*tair + slope_P + elev_mean + cont_area + month.f + wetland + shrub +forest, data = temp,
                na.action = "na.fail")

dredge(full.model)

chosen.model <- lm(ma_mean ~ tair  + month*tair  + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = temp,
                na.action = "na.fail")
summary(chosen.model)
plot(chosen.model)
summ(chosen.model,confint = TRUE, digits = 3)

#save predicted an residual
temp$predicted <- predict(chosen.model)
temp$residuals <- residuals(chosen.model)
hist(temp$residuals)
ggplot(data = temp, aes(x= ma_mean, y=predicted, colour=year)) + geom_point()
ggplot(data = temp, aes(x= ma_mean, y=predicted, colour=Site)) + geom_point()

ggplot(data = temp, aes(x= ma_mean, y=residuals, color=Site)) + geom_point()
ggplot(data = temp, aes(x= day, y=residuals, group=year, color=year)) + geom_point()+ facet_wrap(~Site)
ggplot(data = temp, aes(x= day, y=ma_mean, group=year, color=year)) + geom_point()+ facet_wrap(~Site)

#check out high residuals
resid <- temp %>% filter(residuals > 3)
ggplot(data = resid, aes(x= ma_mean, y=residuals, color=Site)) + geom_point()
ggplot(data = resid, aes(x= ma_mean, y=predicted, color=Site)) + geom_point()

#star <- temp %>% filter(Site=='STAR 171 Lower') 
#ggplot(data=resid, aes(y=ma_mean, x=sampleDate)) + geom_point()



```
### Investigate model by month to see how coefficients behave
June Model
```{r}
June <- temp %>% filter(month==6)
full.modelJ <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = June,
                na.action = "na.fail")

dredge(full.modelJ)

chosen.modelJ <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = June,
                na.action = "na.fail")
summary(chosen.modelJ)
summ(chosen.modelJ,confint = TRUE, digits = 3)
```
### Investigate model by month to see how coefficients behave
July Model
```{r}
July <- temp %>% filter(month==7)
full.modelJY <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = July,
                na.action = "na.fail")

dredge(full.modelJY)

chosen.modelJY <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = July,
                na.action = "na.fail")
summary(chosen.modelJY)
summ(chosen.modelJY, confint = TRUE, digits = 3)
```

### Investigate model by month to see how coefficients behave
August Model
```{r}
August <- temp %>% filter(month==8)
full.modelA <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = August,
                na.action = "na.fail")

dredge(full.modelA)

chosen.modelA <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = August,
                na.action = "na.fail")
summary(chosen.modelA)
summ(chosen.modelA, confint = TRUE, digits = 3)
```

### Investigate model by month to see how coefficients behave
September Model
```{r}
Sept <- temp %>% filter(month==9)
full.modelS <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = Sept,
                na.action = "na.fail")

dredge(full.modelS)

chosen.modelS <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = Sept,
                na.action = "na.fail")
summary(chosen.modelS)
summ(chosen.modelS, confint = TRUE, digits = 3)
```




