---
title: "2_model exploration"
author: "Leslie Jones"
date: "1/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)

library(ggpmisc)
library(lubridate)
library(rgdal)
library(zoo)
library(sf)
library(corrplot)
library(MuMIn)
# library(ape)
library(readxl)
#library(VIF)
library(gridExtra)
library(tidyverse)
#library(jtools)
```

# Database for Anchor Model
Load data frame from 1_Create model data frame.Rmd. Use the full data frame that has 8-day daymet and also lst for comparisons. We will be using a different one for modeling that has already been saved: 3-day dayment and 5-day summed precipitation.

```{r}
load(file = "output/temp.dm2.Rdata")

```


## Compare LST versus daymet air temperatures

See correlation between lst and daymet.
Test models of stream temp with each. This is for the 8day daymet air temps.

```{r}

cor(temp.dm2$lst, temp.dm2$tair, use = "pairwise.complete.obs")

temp.dm2 %>% 
  filter(!is.na(lst)) %>% 
  mutate(diff = lst - tair) %>% 
  summarize(mean(diff), sd(diff))

temp.dm2 %>% 
  ggplot() +
  geom_point(aes(x = tair, y = lst)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)


```

Look at correlations between each air temperature covariate with stream temperature and also predictors of stream temperature.

```{r}
temp.dm2 %>% 
  filter(!is.na(lst), !is.na(ma_mean)) %>% 
  summarize(clst = cor(lst, ma_mean),
            cdm8 = cor(tair, ma_mean))


lm1 <- lm(ma_mean ~ tair, data = temp.dm2 %>% filter(!is.na(lst)))
lm2 <- lm(ma_mean ~ lst, data = temp.dm2 %>% filter(!is.na(lst)))
AIC(lm1, lm2)

summary(lm1)
summary(lm2)

```

Visualize differences

```{r}

temp.dm2 %>% 
  filter(!is.na(lst), !is.na(ma_mean)) %>% 
  select(Site, sampleDate, month, meanT, tair, lst) %>% 
  pivot_longer(cols = tair:lst, names_to = "air_var", values_to = "value") %>% 
  mutate(air_varf = factor(air_var, labels = c("LST", "8-day Daymet"))) %>% 
  ggplot(aes(x = value, y = meanT, color = as.factor(month))) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~air_varf, nrow = 3) +
  stat_poly_eq(formula = y ~ x, parse = TRUE) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(x = "Air Temperature Covariates ", y = "Mean ")

```



## Compare 3-day and 8-day daymet air temperatures

This comparison is slightly different from the one above because now we are using all daily stream temperatures. 

```{r}
temp.dm2 %>% 
  summarize(cdm8 = cor(tair, meanT),
            cdm3 = cor(tair3, meanT))


lm1 <- lm(meanT ~ tair3, data = temp.dm2)
lm2 <- lm(meanT ~ tair, data = temp.dm2)
AIC(lm1, lm2)

summary(lm1)
summary(lm2)


g1 <- ggplot(data = temp, aes(x = tair3, y = meanT)) +
  geom_point() +
  geom_smooth() +
  geom_abline(aes(slope = 1, intercept = 0))

g2 <- ggplot(data = temp, aes(x = tair, y = meanT)) +
  geom_point() +
  geom_smooth() +
  geom_abline(aes(slope = 1, intercept = 0))

grid.arrange(g1,g2)

```

## Explore correlations between the covariates

Spatial covariates first.


```{r}

temp.dm2 %>% 
  distinct(reach_slope, reach_length, Forest, Shrub, Wetland, Riparian, 
           rca_elev_min, rca_elev_max, rca_elev_mn, 
           ca_elev_mn, cont_area) %>% 
  cor(.) %>% 
  corrplot.mixed(.)

#list to match deshka model
temp.dm2 %>% 
  distinct(reach_slope, Forest, Shrub, Wetland, 
           rca_elev_mn, ca_elev_mn, cont_area) %>% 
  cor(.) %>% 
  corrplot.mixed(.)

```




## Exploratory Graphs 
Nothing below has been run/updated, 8/24/20.

### Correlation between LST and air temperature by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(lst2, tair, use = "pairwise.complete.obs"))
```

### correlation between stream temp and air temperature by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(ma_mean, tair ))
```

### correlation and plot between stream temp and LST by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(ma_mean, lst2, use = "pairwise.complete.obs"))

temp.dm %>% 
  ggplot() +
  geom_point(aes(x = lst2, y = ma_mean, color = as.factor(month))) +
  facet_wrap(~Site) 
```

### plot air temperature and LST by month
```{r, echo=FALSE, warnings = FALSE}
temp.dm %>% 
  ggplot() +
  geom_point(aes(x = tair, y = lst2)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)

```
### plot air temperature and stream temperature by month

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
temp.dm %>% 
  ggplot() +
  geom_point(aes(x = tair, y = ma_mean)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)

month5 <- filter(temp.dm,month==5)   
  ggplot() +
  geom_point(data=month5, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
  month6 <- filter(temp.dm,month==6)   
  ggplot() +
  geom_point(data=month6, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
  month7 <- filter(temp.dm,month==7)   
  ggplot() +
  geom_point(data=month7, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
   month8 <- filter(temp.dm,month==8)   
  ggplot() +
  geom_point(data=month8, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")
  
  month9 <- filter(temp.dm,month==9)   
  ggplot() +
  geom_point(data=month9, aes(x = day, y = ma_mean, color=year)) +
  scale_y_continuous(limits = c(0, 31)) + 
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~Site, scales="free")

```


## Exploratory Modeling

Try out sets of models.

```{r, echo=FALSE}

corrplot.mixed(cor(temp %>% select(ma_mean, tair, prcp, swe, slope_P, elev_mean, cont_area, wetland, forest, shrub)))

#model1 <- lm(ma_mean ~ tair, data=temp.mod)
#model2 <- lm(ma_mean ~ tair + elev_mean, data=temp.mod)
#model3 <- lm(ma_mean ~ tair + elev_mean + slope_P, data=temp.mod)
#model4 <- lm(ma_mean ~ tair + elev_mean + slope_P + cont_area, data=temp.mod)
#model5 <- lm(ma_mean ~ tair + elev_mean + slope_P + cont_area + prcp, data=temp.mod)
#AIC(model1, model2, model3, model4, model5)

#make month a factor
month.f <- factor(temp$month)
is.factor(temp$month.f)

full.model <- lm(ma_mean ~ tair + month*tair + slope_P + elev_mean + cont_area + month.f + wetland + shrub +forest, data = temp,
                na.action = "na.fail")

dredge(full.model)

chosen.model <- lm(ma_mean ~ tair  + month*tair  + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = temp,
                na.action = "na.fail")
summary(chosen.model)
plot(chosen.model)
summ(chosen.model,confint = TRUE, digits = 3)

#save predicted an residual
temp$predicted <- predict(chosen.model)
temp$residuals <- residuals(chosen.model)
hist(temp$residuals)
ggplot(data = temp, aes(x= ma_mean, y=predicted, colour=year)) + geom_point()
ggplot(data = temp, aes(x= ma_mean, y=predicted, colour=Site)) + geom_point()

ggplot(data = temp, aes(x= ma_mean, y=residuals, color=Site)) + geom_point()
ggplot(data = temp, aes(x= day, y=residuals, group=year, color=year)) + geom_point()+ facet_wrap(~Site)
ggplot(data = temp, aes(x= day, y=ma_mean, group=year, color=year)) + geom_point()+ facet_wrap(~Site)

#check out high residuals
resid <- temp %>% filter(residuals > 3)
ggplot(data = resid, aes(x= ma_mean, y=residuals, color=Site)) + geom_point()
ggplot(data = resid, aes(x= ma_mean, y=predicted, color=Site)) + geom_point()

#star <- temp %>% filter(Site=='STAR 171 Lower') 
#ggplot(data=resid, aes(y=ma_mean, x=sampleDate)) + geom_point()



```
### Investigate model by month to see how coefficients behave
June Model
```{r}
June <- temp %>% filter(month==6)
full.modelJ <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = June,
                na.action = "na.fail")

dredge(full.modelJ)

chosen.modelJ <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = June,
                na.action = "na.fail")
summary(chosen.modelJ)
summ(chosen.modelJ,confint = TRUE, digits = 3)
```
### Investigate model by month to see how coefficients behave
July Model
```{r}
July <- temp %>% filter(month==7)
full.modelJY <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = July,
                na.action = "na.fail")

dredge(full.modelJY)

chosen.modelJY <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = July,
                na.action = "na.fail")
summary(chosen.modelJY)
summ(chosen.modelJY, confint = TRUE, digits = 3)
```

### Investigate model by month to see how coefficients behave
August Model
```{r}
August <- temp %>% filter(month==8)
full.modelA <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = August,
                na.action = "na.fail")

dredge(full.modelA)

chosen.modelA <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = August,
                na.action = "na.fail")
summary(chosen.modelA)
summ(chosen.modelA, confint = TRUE, digits = 3)
```

### Investigate model by month to see how coefficients behave
September Model
```{r}
Sept <- temp %>% filter(month==9)
full.modelS <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + wetland + shrub +forest, data = Sept,
                na.action = "na.fail")

dredge(full.modelS)

chosen.modelS <- lm(ma_mean ~ tair + slope_P + elev_mean + cont_area + forest + shrub + wetland  , data = Sept,
                na.action = "na.fail")
summary(chosen.modelS)
summ(chosen.modelS, confint = TRUE, digits = 3)
```




