---
title: "2_model exploration"
author: "Leslie Jones"
date: "1/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)

library(lubridate)
library(rgdal)
library(zoo)
library(sf)
library(corrplot)
library(MuMIn)
library(ape)
library(readxl)
library(VIF)
library(tidyverse)
```

## Database for Anchor Model
Load data frame from 1_data analysis.Rmd

```{r}
load(file = "output/temp.dm")
#remove April
temp.dm <- filter(temp.dm, month > 4)


```

## Explore LST and daymet as covariates

correlation between LST and air temperature by month
```{r}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(lst2, tair, use = "pairwise.complete.obs"))
```

correlation between stream temp and air temperature by month
```{r}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(ma_mean, tair ))
```

correlation between stream temp and LST by month
```{r}
temp.dm %>%
  group_by(month) %>%
  summarise(cor = cor(ma_mean, lst2, use = "pairwise.complete.obs"))
```

plot air temperature and LST by month
```{r}
temp.dm %>% 
  ggplot() +
  geom_point(aes(x = tair, y = lst2)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)

```
plot air temperature and stream temperature by month
```{r}
temp.dm %>% 
  ggplot() +
  geom_point(aes(x = tair, y = ma_mean)) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 25)) +
  facet_wrap(~month)
```

plot LST versus stream temperature.

```{r}
temp.dm %>% 
  ggplot() +
  geom_point(aes(x = lst2, y = ma_mean, color = as.factor(month))) +
  facet_wrap(~Site) 

```


## Modeling

Try out sets of models.

```{r}

corrplot.mixed(cor(temp.dm %>% select(tair, prcp, swe, slope_P, elev_mean, cont_area)))

model1 <- lm(ma_mean ~ tair, data=temp.dm)
model2 <- lm(ma_mean ~ tair + elev_mean, data=temp.dm)
model3 <- lm(ma_mean ~ tair + elev_mean + slope_P, data=temp.dm)
model4 <- lm(ma_mean ~ tair + elev_mean + slope_P + cont_area, data=temp.dm)
model5 <- lm(ma_mean ~ tair + elev_mean + slope_P + cont_area + prcp, data=temp.dm)
AIC(model1, model2, model3, model4, model5)

#make month a factor
month.f <- factor(temp.dm$month)
is.factor(temp.dm$month.f)

full.model <- lm(ma_mean ~ tair + prcp + swe + slope_P + elev_mean + cont_area + month.f, data = temp.dm,
                 na.action = "na.fail")

summary(full.model)
dredge(full.model)

summary(lm)

```

