---
title: "5_spatial_predictions"
author: "Leslie"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tidyverse)
library(rgdal)
library(lubridate)
library(gridExtra)
```

# Predictions data frame

## Spatial covariates

Read in spatial covariates associated with each RCA. These are stored in two separate feature classes.


```{r}
fgdb <- "Spatial_data/Anchor//Anchor.gdb"
# List all feature classes in a file geodatabase
subset(ogrDrivers(), grepl("GDB", name))
fc_list <- ogrListLayers(fgdb)
print(fc_list)
# Read the feature class
rca_reach <- sf::st_read(dsn = fgdb, layer = "anch_rca_reaches_attributed_06022020") %>% as.data.frame
rcas <- sf::st_read(dsn = fgdb, layer = "anchor_rcas_attributed_06022020") %>% as.data.frame
```

Create a new data frame that will be used for predictions. Use the flowacc value and convert to square kilometers. (The cont_area is the flow accumulation multiplied by 25 and converted from square meters to square kilometers.)

```{r}
new_data_sp <- rca_reach %>% 
  select(rca_id = reach_id, reach_slope, reach_length, Forest, Shrub, Wetland, Riparian) 

new_data_sp <- new_data_sp %>% 
  left_join(rcas %>% select(rca_id:ca_elev_mn)) %>% 
  mutate(cont_area = flowacc * 25 / 1000000)
```

## NOT DONE: Filter on salmon rcas

Note (2/27/20): Decided to predict for all rcas so we can compare salmon vs potential salmon habitat in the future.

Filter salmon rcas only. This is the dataframe that we will predict on. Look at distribution of covariates that will predict over compared to rcas used in model parameterization.

```{r}
salmon_rcas <- new_data_sp %>% 
  mutate(salmon = K_p + CH_p + P_p + CO_p + S_p) %>% 
  filter(salmon > 0) 
nrow(salmon_rcas)
salmon_rcas %>% 
  select(rca_id, rca_elev_mn, ca_elev_mn, cont_area, reach_slope) %>% 
  cov(.)
hist(salmon_rcas$ca_elev_mn)
hist(salmon_rcas$rca_elev_mn)
hist(salmon_rcas$reach_slope)
hist(salmon_rcas$cont_area)
```

## Add Daymet data to new_data data frame

Read in daymet data. Get 3-day air temperature and 5-day precip. No longer filtering by rcas for salmon, these are now all rcas.

```{r}
fgdb <- "Data/Spatial_data/Anchor/KFHP/Geodatabases/anchor_DAYMET.gdb"
subset(ogrDrivers(), grepl("GDB", name))
fc_list <- ogrListLayers(fgdb)
print(fc_list)
dm3day <- sf::st_read(dsn = fgdb, layer = "tair3_anchor_all")

dm3day <- dm3day %>% 
  mutate(date = as.Date(date)) %>% 
  select(-Field1) %>% 
  mutate(date_day3 = date + 1)

new_data_cl <- dm3day %>%
  filter(month(date_day3) %in% 6:9) %>% 
  select(-date) 

summary(new_data_cl)

new_data_cl %>% filter(tair3 < 0)
```

Daymet precipitation.

```{r}

prcp5day <- sf::st_read(dsn = fgdb, layer = "prcp5day")
prcp5day <- prcp5day %>% 
  mutate(date = as.Date(date)) %>% 
  select(-Field1) %>% 
  mutate(date_day5 = date + 4)

prcp5day <- prcp5day %>% select(-date)

new_data_cl <- new_data_cl %>% 
  left_join(prcp5day, by = c("rca_id" = "rca_id", "date_day3" = "date_day5"))

```


Merge climate and spatial data frames. Join on the climate data so that spatial covariates are repeated across all dates.

```{r}
new_data <- left_join(new_data_cl %>% rename(date = date_day3), new_data_sp)
```

Add julian date so that the spring model and fall models are predicting only within the range of dates that were used to build the models.

```{r}
new_data <- new_data %>% 
  mutate(day = as.numeric(format(date, "%j")),
         year = year(date))
```


Add in April 1 swe.

```{r}
sweA1 <- read_csv("Data/sweA1_anchor.csv")

sweA1 <- sweA1 %>% 
  mutate(year = year(date)) %>% 
  select(-X1, -date)

new_data <- new_data %>% 
  left_join(sweA1)
```


# Predictions over new data frame

Load models from scrip 3.

```{r}
load(file = "output/spring.lm.Rdata")
load(file = "output/fall.lm.Rdata")
```

Make predictions for spring

```{r}
new_data_spring <- new_data %>% 
  filter(day < 202)
new_data_spring <- new_data_spring %>% 
  mutate(preds = predict(spring.lm, newdata = new_data_spring))
```

Make predictions for fall.

```{r}
new_data_fall <- new_data %>% 
  filter(day > 201)
new_data_fall <- new_data_fall %>% 
  mutate(preds = predict(fall.lm, newdata = new_data_fall))
```


Combine.

```{r}
preds <- bind_rows(new_data_spring, new_data_fall)
```

Save predictions data frame.

```{r}
save(preds, file = "output/preds.Rdata")
```






