---
title: "6_Thermal Niche"
author: "Leslie Jones"
date: "2/10/2020"
output:
  html_document:
    fig_width: 4
    fig_height: 4
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(tidyverse)
library(rgdal)
library(lubridate)
library(gridExtra)
library(arsenal)
library(knitr)
library(magick)
library(gganimate)
library(transformr)
library(kableExtra)
```

# Thermal niche summaries for Chinook Salmon
Spawning habitat will be summarized for July, August, and September, whereas rearing habitat is summarized for June, July, and August.

```{r}
#load chinook rearing and spawning datasets
load(file = "output/K_r.Rdata")
load(file = "output/K_s.Rdata")
load(file = "output/K_p.Rdata")
```


##Interannual Variability of all Chinook habitat
To summarize interannual variability of Chinook habitat we averaged temperatures from June 1 - September 30 for each year and plotted the results. Stanadard deviation of yearly averages represent the spatial variability of RCAs. 
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
#How about looking at interannual variability
interannual <- K_p %>% group_by(year) %>% mutate(yr_mn=mean(preds), yr_sd=sd(preds)) 

ggplot(data=interannual) +
  geom_line(aes(x = year, y = yr_mn))  + geom_smooth(aes(x = year, y = yr_mn),method = "lm") +
  labs(x="Year", y="Stream Temperature")+
  ggtitle("Mean annual stream temperatures for Chinook habitat") 

ggplot(data=interannual) +
  geom_line(aes(x = year, y = yr_sd)) + geom_smooth(aes(x = year, y = yr_sd),method = "lm") +
  labs(x="Year", y="Standard Deviation of Stream Temperatures")+
  ggtitle("Variability of mean annual temperatures among RCAs") 

#explore another table method
tab <- K_p %>% group_by(year) %>% summarize(mean = mean(preds), std=sd(preds))
tab %>% knitr::kable(caption = "Mean annual temperatures for Chinook habitat", digits = 2) %>%     kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)

```
##Interannual variability of monthly temperatures
Average monthly temperatures were summ
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
#How about looking at interannual variability
intermonth <- K_p %>% group_by(year, month) %>% mutate(mth_mn=mean(preds), mth_sd=sd(preds)) 

intermonth$month2 <- factor(intermonth$month, labels=c("June", "July", "August", "September"))

ggplot(data=intermonth) +
  geom_line(aes(x = year, y =mth_mn))  +
  labs(x="Year", y="Average Stream Temperature")+
  ggtitle("Mean monthly stream temperatures for Chinook presence habitat") +
   facet_grid(~month2) 

ggplot(data=intermonth) +
  geom_line(aes(x = year, y = mth_sd))  + 
  labs(x="Year", y="SD of Stream Temperature") + 
   ggtitle("Variability of mean monthly stream temperatures among RCAs") +
   facet_grid(~month2) 

#explore another table method
tab1 <- K_p %>% group_by(year, month) %>% summarize(mean = mean(preds), std=sd(preds))

Junt <- tab1 %>% filter(month==6)
Junt %>% knitr::kable(caption = "Mean June temperatures for Chinook habitat", digits = 2) %>%     kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)

Jult <- tab1 %>% filter(month==7)
Jult %>% knitr::kable(caption = "Mean July temperatures for Chinook habitat", digits = 2) %>%     kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)

Aug <- tab1 %>% filter(month==8)
Aug %>% knitr::kable(caption = "Mean August temperatures for Chinook habitat", digits = 2) %>%     kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)

Sep <- tab1 %>% filter(month==9)
Sep %>% knitr::kable(caption = "Mean September temperatures for Chinook habitat", digits = 2) %>%     kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
```



#Mean daily temperatures across 38 year period

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}

#Chinook presence
mean_p <- K_p %>% group_by(day) %>% summarize(mean_day = mean(preds), sd_p = sd(preds)) 

ggplot(data=mean_p) +
  geom_smooth(aes(x = day, y = mean_day), show.legend = FALSE) + labs(x= "Day of Year", y="Stream Temperature") +
  ggtitle("Mean daily predictions 1980-2018 for Chinook presence habitat") 

ggsave("Anchor_chinook_plots/Chinook Presence: Mean daily temperature 1980-2018.jpeg",  plot=last_plot())

#Chinook spawning
mean_s <- K_s %>% group_by(day) %>% 
  summarize(mean_day = mean(preds), sd_p = sd(preds))

ggplot(data=mean_s) +
  geom_smooth(aes(x = day, y = mean_day), show.legend = FALSE) + labs(x= "Day of Year", y="Stream Temperature") +
  ggtitle("Mean daily predictions 1980-2018 for Chinook spawning") 

ggsave("Anchor_chinook_plots/Chinook Spawning: Mean daily temperature 1980-2018.jpeg", plot=last_plot())

#Chinook rearing
mean_r <- K_r %>% group_by(day) %>%
  summarize(mean_day = mean(preds), sd_p = sd(preds))
  
ggplot(data=mean_r) +
  geom_smooth(aes(x = day, y = mean_day), show.legend = FALSE) + labs(x= "Day of Year", y="Stream Temperature") +
  ggtitle("Mean daily predictions 1980-2018 for Chinook rearing") 

ggsave("Anchor_chinook_plots/Chinook Rearing: Mean daily temperature 1980-2018.jpeg", plot=last_plot())


```
#histograms of percent habitat for each degree celcius - Chinook Presence habitat
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
hist <- K_p %>% group_by(rca_id, month, decade) %>% summarize(mth_mn=mean(preds)) %>% 
  mutate(bin=case_when(mth_mn>=1 & mth_mn<2 ~ 1, mth_mn>=2 & mth_mn<3 ~ 2, mth_mn>=3 & mth_mn<4 ~ 3,
                                     mth_mn>=4 & mth_mn<5 ~ 4, mth_mn>=5 & mth_mn<6 ~ 5, mth_mn>=6 & mth_mn<7 ~ 6,
                                     mth_mn>=7 & mth_mn<8 ~ 7, mth_mn>=8 & mth_mn<9 ~ 8, mth_mn>=9 & mth_mn<10 ~ 9,
                                     mth_mn>=10 & mth_mn<11 ~ 10, mth_mn>=11 & mth_mn<12 ~ 11, mth_mn>=12 & mth_mn<13 ~ 12,
                                     mth_mn>=13 & mth_mn<14 ~ 13, mth_mn>=14 & mth_mn<15 ~ 14))

#calculate total chinook presence habitat = 321262.5
total <- K_p %>%  filter(year==1980 & day==220) %>% summarize(total_hab=sum(reach_length))

len <- K_p %>% distinct(rca_id, reach_length) 

hist2 <- left_join(hist, len, by="rca_id") %>% group_by(decade, month, bin) %>% summarize(bin_total=sum(reach_length)) %>% 
        mutate(percent=(bin_total/321262.5)*100) 


#June
jun <- hist2 %>% filter(month==6)

ggplot(data=jun) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + labs(x="Temperature", y="Percent total habitat", title="Mean June Stream Temperatures")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

#July
jul <- hist2 %>% filter(month==7)

ggplot(data=jul) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + labs(x="Temperature", y="Percent total habitat", title="Mean July Stream Temperatures")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

#aug
aug <- hist2 %>% filter(month==8)

ggplot(data=aug) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + labs(x="Temperature", y="Percent total habitat", title="Mean August Stream Temperatures")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

#sep
sep <- hist2 %>% filter(month==9)

ggplot(data=sep) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + labs(x="Temperature", y="Percent total habitat", title="Mean September Stream Temperatures")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))


```
#histograms of percent habitat for each degree celcius - Chinook Spawning habitat
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
hist_s <- K_s %>% group_by(rca_id, month, decade) %>% summarize(mth_mn=mean(preds)) %>% 
  mutate(bin=case_when(mth_mn>=1 & mth_mn<2 ~ 1, mth_mn>=2 & mth_mn<3 ~ 2, mth_mn>=3 & mth_mn<4 ~ 3,
                                     mth_mn>=4 & mth_mn<5 ~ 4, mth_mn>=5 & mth_mn<6 ~ 5, mth_mn>=6 & mth_mn<7 ~ 6,
                                     mth_mn>=7 & mth_mn<8 ~ 7, mth_mn>=8 & mth_mn<9 ~ 8, mth_mn>=9 & mth_mn<10 ~ 9,
                                     mth_mn>=10 & mth_mn<11 ~ 10, mth_mn>=11 & mth_mn<12 ~ 11, mth_mn>=12 & mth_mn<13 ~ 12,
                                     mth_mn>=13 & mth_mn<14 ~ 13, mth_mn>=14 & mth_mn<15 ~ 14))

#calculate total chinook spawning habitat = 	284418.7
total_s <- K_s %>%  filter(year==1980 & day==220) %>% summarize(total_hab=sum(reach_length))

len_s <- K_s %>% distinct(rca_id, reach_length) 

hist2_s <- left_join(hist_s, len_s, by="rca_id") %>% group_by(decade, month, bin) %>% 
    summarize(bin_total=sum(reach_length)) %>% 
        mutate(percent=(bin_total/284418.7)*100) 

#July
jul_s <- hist2_s %>% filter(month==7)

ggplot(data=jul_s) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + 
  labs(x="Temperature", y="Percent total habitat", title="Mean July Stream Temperatures for Chinook Spawning habitat")+
  scale_fill_discrete(name="Decade") + scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

#aug
aug_s <- hist2_s %>% filter(month==8)

ggplot(data=aug_s) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + 
  labs(x="Temperature", y="Percent total habitat", title="Mean August Stream Temperatures for Chinook Spawning habitat")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

#sep
sep_s <- hist2_s %>% filter(month==9)

ggplot(data=sep_s) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + 
  labs(x="Temperature", y="Percent total habitat", title="Mean September Stream Temperatures for Chinook spawning habitat")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))
```

#histograms of percent habitat for each degree celcius - Chinook Rearing habitat
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
hist_r <- K_r %>% group_by(rca_id, month, decade) %>% summarize(mth_mn=mean(preds)) %>% 
  mutate(bin=case_when(mth_mn>=1 & mth_mn<2 ~ 1, mth_mn>=2 & mth_mn<3 ~ 2, mth_mn>=3 & mth_mn<4 ~ 3,
                                     mth_mn>=4 & mth_mn<5 ~ 4, mth_mn>=5 & mth_mn<6 ~ 5, mth_mn>=6 & mth_mn<7 ~ 6,
                                     mth_mn>=7 & mth_mn<8 ~ 7, mth_mn>=8 & mth_mn<9 ~ 8, mth_mn>=9 & mth_mn<10 ~ 9,
                                     mth_mn>=10 & mth_mn<11 ~ 10, mth_mn>=11 & mth_mn<12 ~ 11, mth_mn>=12 & mth_mn<13 ~ 12,
                                     mth_mn>=13 & mth_mn<14 ~ 13, mth_mn>=14 & mth_mn<15 ~ 14))

#calculate total chinook rearing habitat = 	176816.9
total_r <- K_r %>%  filter(year==1980 & day==220) %>% summarize(total_hab=sum(reach_length))

len_r <- K_r %>% distinct(rca_id, reach_length) 

hist2_r <- left_join(hist_r, len_r, by="rca_id") %>% group_by(decade, month, bin) %>% 
  summarize(bin_total=sum(reach_length)) %>% 
        mutate(percent=(bin_total/176816.9)*100)

#June
jun_r <- hist2_r %>% filter(month==6)

ggplot(data=jun_r) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + 
  labs(x="Temperature", y="Percent total habitat", title="Mean June Stream Temperatures for Chinook   Rearing habitat")+  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

#July
jul_r <- hist2_r %>% filter(month==7)

ggplot(data=jul_r) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + 
  labs(x="Temperature", y="Percent total habitat", title="Mean July Stream Temperatures for Chinook Rearing habitat")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

#aug
aug_r <- hist2_r %>% filter(month==8)

ggplot(data=aug_r) + geom_bar(aes(x=bin, y=percent, fill=as.factor(decade)), stat = "identity")+
  facet_wrap(~decade) + 
  labs(x="Temperature", y="Percent total habitat", title="Mean August Stream Temperatures for Chinook Rearing habitat")+
  scale_fill_discrete(name="Decade")+ scale_x_continuous(breaks = c(2,4,6,8,10,12,14))

```

##Daily mean tempertures for each year (1980-2018)
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}

# Summarize mean and sd of predictions by day and year - not sure how to retain variables when using the summarize function?? so create decade variable within

all_years <- K_p %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds))

ggplot(data=all_years) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), se=FALSE, show.legend = FALSE) + 
  ggtitle("Annual mean of daily predictions") 

ggsave("Anchor_chinook_plots/Annual means.jpeg", plot=last_plot())

decade <- K_p %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) %>% 
  mutate(decade=case_when(year< 1990 ~ 1980, year >= 1990 & year < 2000 ~ 1990, 
                          year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010))

ggplot(data=decade) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), show.legend = FALSE) + facet_wrap(~decade) + 
  ggtitle("Annual mean of daily predictions by decade") 

ggsave("Anchor_chinook_plots/Annual means by decade.jpeg", plot=last_plot())

decade_noSE <- K_p %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds))  %>% 
  mutate(decade=case_when(year< 1990 ~ 1980, year >= 1990 & year < 2000 ~ 1990, 
                          year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010))

ggplot(data=decade_noSE) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), se=FALSE, show.legend = FALSE) + facet_wrap(~decade) +
  ggtitle("Annual mean of daily predictions by decade no SE") 

ggsave("Anchor_chinook_plots/Annual means by decade no SE.jpeg", plot=last_plot())
```


```{r}
#Animate by Year using gganimate
#https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautiful-animation-in-r/
#shadow fade with old years in background
theme_set(theme_bw())
ani <- ggplot(all_years, aes(x = day, y = mean_yr, color = as.factor(year))) +
  geom_smooth(se=FALSE, show.legend = FALSE) + 
  scale_size(range = c(2, 12)) +
  transition_states(year, transition_length = 1, state_length = 1) + shadow_mark(alpha=0.3, size=0.7) + 
  transition_time (year) + 
  labs(x = "Day of Year", y = "Daily mean Stream Temperature", title = "Year: {as.integer(frame_time)}")
  

animate(ani, duration=50, fps=1)

#try the "Let data gradually appear" with month time series
K_p2 <- preds %>% mutate(month=month(date), day2=(day(date))) %>% mutate(decade=case_when(year< 1990 ~ 1980, year >= 1990 & year < 2000 ~ 1990, year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010)) %>% filter(K_p == 1)  

all_years2 <- K_p2 %>% group_by(month, day2) %>% 
  summarize(mean_mth = mean(preds))

ani2 <- ggplot(all_years2, aes(x = day2, y = mean_mth, color=factor(month))) +
  geom_line() +
  labs(x = "Day of Month", y = "Stream Temperature", title = "Average Daily Temperatures 1980-2018") 

ani2 + transition_reveal(day2)

```

##Animate months by decade
```{r}
all_years22 <- K_p2 %>% group_by(month, day2, decade) %>% 
  summarize(mean_mth_dec = mean(preds))

ggplot(all_years22, aes(x = day2, y = mean_mth_dec, color=factor(month))) +
  geom_line() +  facet_wrap(~decade) +
  labs(x = "Day of Month", y = "Stream Temperature", title = "Average Daily Temperatures by decade") 

ggsave("Anchor_chinook_plots/Average Daily Temps Animated.jpeg", plot=last_plot())

ani3 <- ggplot(all_years22, aes(x = day2, y = mean_mth_dec, color=factor(month))) +
  geom_line() + 
  labs(x = "Day of Month", y = "Stream Temperature", title = "Average Daily Temperatures by decade") 

ani3 + transition_reveal(day2) + facet_wrap(~decade)
```



##Summarize predictions by decades and plot
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
#Summarize mean and sd of preds by day and decade
pred_decade <- K_p %>% group_by(decade, day) %>% 
  summarize(mean_dec = mean(preds), sd_dec = sd(preds))

ggplot(data=pred_decade) +
  geom_smooth(aes(x = day, y = mean_dec, color = as.factor(decade))) +
  ggtitle("Decadal mean of daily predictions") 

ggsave("Anchor_chinook_plots/Decadal mean of daily predictions.jpeg", plot=last_plot())

#Summarize mean and sd of preds by day and decade - no SE
pred_decade_noSE <- K_p %>% group_by(decade, day) %>% 
  summarize(mean_dec = mean(preds), sd_dec = sd(preds)) 

ggplot(data=pred_decade_noSE) +
  geom_smooth(aes(x = day, y = mean_dec, color = as.factor(decade)), se=FALSE) +
  ggtitle("Decadal mean of daily predictions no SE") 

ggsave("Anchor_chinook_plots/Decadal mean of daily predictions no SE.jpeg", plot=last_plot())


#Summarize mean and sd of preds by day and decade - facet by decade
pred_decade2 <- K_p %>% group_by(decade, day) %>% 
  summarize(mean_dec = mean(preds), sd_dec = sd(preds)) 

ggplot(data=pred_decade2) +
  geom_smooth(aes(x = day, y = mean_dec)) + facet_wrap(~decade)+
  ggtitle("Decadal mean of daily predictions") 

ggsave("Anchor_chinook_plots/Annual means by decade facet by decade.jpeg", plot=last_plot())

```


##Calculate and plot Temperature Anomolies!

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
#Calculate annual mean 1980-2018 then yearly differences between 38 year mean. 
#calculate average daily predictions over entire 38 year period 
pred_anom1 <- K_p %>% 
  group_by(day) %>% 
  summarize(mean38 = mean(preds), sd38 = sd(preds)) 

#calculate difference between 38 year average and daily prediction dataframe
pred_anom2 <- left_join(preds, pred_anom1) %>% mutate(anomoly=(mean38-preds))

#summarize anomolies by year and plot by decade
anomoly <- pred_anom2 %>% group_by(day, year) %>% 
  summarize(anom_yr = mean(anomoly)) %>% mutate(decade=case_when(year< 1990 ~ 1980, 
  year >= 1990 & year < 2000 ~ 1990, year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010)) 

ggplot(data=anomoly) +
  geom_smooth(aes(x = day, y = anom_yr, color = as.factor(year)), show.legend = FALSE) + facet_wrap(~decade) +
  ggtitle("Anomolies by day and year faceted by decade") 

ggsave("Anchor_chinook_plots/Anomolies by day and year faceted by decade.jpeg", plot=last_plot())


#summarize anomolies by year and plot by decade - no SE
anomoly_noSE <- pred_anom2 %>% group_by(day, year) %>% 
  summarize(anom_yr = mean(anomoly)) %>% mutate(decade=case_when(year< 1990 ~ 1980, 
  year >= 1990 & year < 2000 ~ 1990, year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010)) 

ggplot(data=anomoly_noSE) +
  geom_smooth(aes(x = day, y = anom_yr, color = as.factor(year)), se=FALSE, show.legend = FALSE) + facet_wrap(~decade)  +
  ggtitle("Anomolies by day and year faceted by decade") 

ggsave("Anchor_chinook_plots/Anomolies by day and year no SE faceted by decade.jpeg", plot=last_plot())


```
#Filter for spawning time period - Chinook reach their spawning areas between July and September  - DESCENDING LIMB!!!!
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
# Summarize mean and sd of predictions by day and year - not sure how to retain variables when using the summarize function?? so create decade variable within

CH_s <- K_s %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) 

ggplot(data=CH_s) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), se=FALSE, show.legend = FALSE)  +
  ggtitle("Annual mean temperatures of Chinook spawning") 

ggsave("Anchor_chinook_plots/Annual mean temperatures of Chinook spawning.jpeg", plot=last_plot())

CH_s_dec <- K_s %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) %>% mutate(decade=case_when(year< 1990 ~ 1980, 
  year >= 1990 & year < 2000 ~ 1990, year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010)) 

ggplot(data=CH_s_dec) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), show.legend = FALSE) + facet_wrap(~decade) +
  ggtitle("Annual mean temperatures of Chinook spawning faceted by decade") 

ggsave("Anchor_chinook_plots/Annual mean temperatures of Chinook spawning faceted by decade.jpeg", plot=last_plot())

CH_s_dec_noSE <- K_s %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) %>% mutate(decade=case_when(year< 1990 ~ 1980, 
  year >= 1990 & year < 2000 ~ 1990, year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010)) 

ggplot(data=CH_s_dec_noSE) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), se=FALSE, show.legend = FALSE) + facet_wrap(~decade) +
  ggtitle("Annual mean temperatures of Chinook spawning faceted by decade no SE") 

ggsave("Anchor_chinook_plots/Annual mean temperatures of Chinook spawning faceted by decade no SE.jpeg", plot=last_plot())


```
#Filter for core rearing - June-August  
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align="center", fig.width=6, fig.asp=1}
# Summarize mean and sd of predictions by day and year - not sure how to retain variables when using the summarize function?? so create decade variable within

CH_r <- K_r %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) 

ggplot(data=CH_r) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), se=FALSE, show.legend = FALSE)+
  ggtitle("Annual mean temperatures of Chinook rearing") 

ggsave("Anchor_chinook_plots/Annual mean temperatures of Chinook rearing.jpeg", plot=last_plot())

CH_r_dec <- K_r %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) %>% mutate(decade=case_when(year< 1990 ~ 1980, 
  year >= 1990 & year < 2000 ~ 1990, year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010)) 

ggplot(data=CH_r_dec) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), show.legend = FALSE) + facet_wrap(~decade)+
  ggtitle("Annual mean temperatures of Chinook rearing faceted by decade") 

ggsave("Anchor_chinook_plots/Annual mean temperatures of Chinook rearing faceted by decade.jpeg", plot=last_plot())

CH_r_dec_noSE <- K_r %>% group_by(day, year) %>% 
  summarize(mean_yr = mean(preds), sd_yr = sd(preds)) %>% mutate(decade=case_when(year< 1990 ~ 1980, 
  year >= 1990 & year < 2000 ~ 1990, year >=2000 & year < 2010 ~ 2000, year>2009 ~ 2010)) 

ggplot(data=CH_r_dec_noSE) +
  geom_smooth(aes(x = day, y = mean_yr, color = as.factor(year)), se=FALSE, show.legend = FALSE) + facet_wrap(~decade)+
  ggtitle("Annual mean temperatures of Chinook rearing faceted by decade no SE") 

ggsave("Anchor_chinook_plots/Annual mean temperatures of Chinook rearing faceted by decade no SE.jpeg", plot=last_plot())


```
