---
title: "7_App C thermal regimes"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

library(RColorBrewer)
library(sjPlot)
library(gridExtra)
library(tidyverse)
library(sf)
library(readxl)
```

Create data frame for figures that can be saved and imported to Deshka appendix c.

# figure 1 data
```{r}

#read in new file with awc attributes on catchments, these were edited on 9/12/20
cat_spp <- read_excel("Data/anchor_rcas_awc_updated_091220.xls")

readRDS(file = "output/mets.rds")

mets_spp <- left_join(mets, cat_spp, by = c("reach_id" = "rca_id"))

mets_spp <- mets_spp %>% 
  mutate(decade = case_when(year > 1979 & year < 1990 ~ "1980s",
                            year > 1989 & year < 2000 ~ "1990s",
                            year > 1999 & year < 2010 ~ "2000s",
                            year > 2009 & year < 2020 ~ "2010s"))

mets_spp <- mets_spp %>% 
  mutate(MA7d_DAT_jd = as.numeric(MA7d_DAT_jd),
         MxDAT_jd = as.numeric(MxDAT_jd))

saveRDS(mets_spp, file = "output/mets_spp.rds")
```


```{r include = FALSE}
fgdb <- "W:/GIS/Anchor/Anchor_spatial.gdb"

fl <- sf::st_read(dsn = fgdb, layer = "anch_rca_reaches_attributed_03112020")
fl <- st_zm(fl)
fl <- fl %>% st_transform(4326)

cat <- sf::st_read(dsn = fgdb, layer = "anchor_rcas_attributed_06022020")
cat <- st_zm(cat)
cat<- cat %>% st_transform(4326)

#remove incorrect species information from catchments
cat <- cat %>% 
  select(-(K_r:S_p))

fl_spp <- merge(fl, cat_spp %>% select(rca_id, K_s, CO_s, K_r, CO_r), by.x = "reach_id", by.y = "rca_id") 
cat_spp <- merge(cat, cat_spp %>% select(rca_id, K_s, CO_s, K_r, CO_r))

fl_spp_tbl <- fl_spp %>% 
  st_set_geometry(NULL) %>% 
  mutate(LengthKM = reach_length/1000)

fl_spp_tbl %>% 
  summarize(sum(LengthKM))

#get correct length onto met_spp
mets_spp <- fl_spp_tbl %>% 
  select(reach_id, LengthKM) %>% 
  right_join(mets_spp)

plot(fl)
```

Create watershed boundary and save.

```{r anchor_wtd, eval = FALSE}

anchor_wtd <- cat %>% 
  mutate(wtd = "Anchor-Stariski") %>% 
  group_by(wtd) %>% 
  summarize()

plot(anchor_wtd)

saveRDS(anchor_wtd, "output/anchor_wtd.rds")

#also flowlines
saveRDS(fl, file = "output/anchor_fl.rds")

```
```



```{r}
#covert to long form so can facet by species and habitat (color = spp, facet = habitat)
ks <- fl_spp %>% 
  filter(K_s > 0) %>% 
  mutate(Species = "Chinook Salmon",
         Habitat = "Spawning")
kr <- fl_spp %>% 
  filter(K_r > 0) %>% 
  mutate(Species = "Chinook Salmon",
         Habitat = "Rearing")
cs <- fl_spp %>% 
  filter(CO_s > 0) %>% 
  mutate(Species = "Coho Salmon",
         Habitat = "Spawning")
cr <- fl_spp %>% 
  filter(CO_r > 0) %>% 
  mutate(Species = "Coho Salmon",
         Habitat = "Rearing")


fl_spp_anchor <- rbind(ks %>% select(-(K_s:CO_r)),
                   kr %>% select(-(K_s:CO_r)),
                   cs %>% select(-(K_s:CO_r)),
                   cr %>% select(-(K_s:CO_r)))

saveRDS(fl_spp_anchor, file = "output/fl_spp_anchor.rds")
```


```{r}
ggplot() +
  geom_sf(data = fl_spp %>% filter(K_r > 0))
```

# figure 2 data

```{r}
#need a long form of species and life stages. Then can filter and facet as needed.

#get column indicating which species has spawning habitat in each rca
spawn_mets <- mets_spp %>%
  ungroup() %>% 
  select(reach_id:MA7d_DAT_jd, K_s, CO_s, LengthKM, decade) %>% 
  pivot_longer(cols = c("K_s", "CO_s"), names_to = "Species", values_to = "present") %>% 
  filter(present > 0)

#get months into their own column so they can be averaged by decade
spawn_months <- spawn_mets %>% 
  ungroup() %>% 
  select(reach_id:Aug_mn, Species, LengthKM, decade) %>% 
  pivot_longer(cols = June_mn:Aug_mn, names_to = "Month")

#take average of months by decade
spawn_months_dec <- spawn_months %>%
  ungroup() %>% 
  group_by(decade, reach_id, LengthKM, Species, Month) %>% 
  summarize(dec_mn = mean(value)) %>% 
  mutate(dec_mn_cuts = cut(dec_mn, breaks = 3:22),
         Species = factor(Species, levels = c("K_s", "CO_s"), 
                          labels = c("Chinook Salmon", "Coho Salmon")),
         Month = factor(Month, levels = c("June_mn", "July_mn", "Aug_mn"),
                        labels = c("June", "July", "August")))

spawn_months_dec_anchor <- spawn_months_dec
saveRDS(spawn_months_dec_anchor, file = "output/spawn_months_dec_anchor.rds")

#need to save spawn months for summary across all 40 years
spawn_months_anchor <- spawn_months
saveRDS(spawn_months_anchor, file = "output/spawn_months_anchor.rds")

```

```{r}
missing_9deg <- spawn_months_dec %>% 
  ungroup() %>% 
  filter(Month == "September", dec_mn > 9) %>% 
  distinct(reach_id)

spawn_months %>% 
  right_join(missing_9deg) %>% 
  count(year)

spawn_months %>% 
  right_join(missing_9deg) %>% 
  filter(Month == "Sep_mn") %>% 
  ggplot(aes(x = year, y = value, color = factor(reach_id))) +
  geom_line()

spawn_months_dec %>% 
  right_join(missing_9deg) %>% 
  filter(Month == "September") 
```

Rearing data

```{r}
#need a long form of species and life stages. Then can filter and facet as needed.

rear_mets <- mets_spp %>%
  ungroup() %>% 
  select(reach_id:MA7d_DAT_jd, K_r, CO_r, LengthKM, decade) %>% 
  pivot_longer(cols = c("K_r", "CO_r"), names_to = "Species", values_to = "present") %>% 
  filter(present > 0)

rear_months <- rear_mets %>% 
  ungroup() %>% 
  select(reach_id:Aug_mn, Species, LengthKM, decade) %>% 
  pivot_longer(cols = June_mn:Aug_mn, names_to = "Month")

rear_months_dec <- rear_months %>%
  ungroup() %>% 
  group_by(decade, reach_id, LengthKM, Species, Month) %>% 
  summarize(dec_mn = mean(value)) %>% 
  mutate(dec_mn_cuts = cut(dec_mn, breaks = 1:13),
         Species = factor(Species, levels = c("K_r", "CO_r"), 
                          labels = c("Chinook Salmon", "Coho Salmon")),
         Month = factor(Month, levels = c("June_mn", "July_mn", "Aug_mn"),
                        labels = c("June", "July", "August")))


rear_months_dec_anchor <- rear_months_dec
saveRDS(rear_months_dec_anchor, file = "output/rear_months_dec_anchor.rds")

#need monthly summaries to get thermal regimes across historical period
rear_months_anchor <- rear_months
saveRDS(rear_months_anchor, file = "output/rear_months_anchor.rds")

```

potential pike habitat

```{r}
mets_spp %>% 
  select(decade, reach_id, LengthKM, July_mn) %>% 
  group_by(decade, reach_id, LengthKM) %>% 
  summarize(july = mean(July_mn)) %>% 
  ungroup() %>% 
  filter(decade == "2010s") %>% 
  summarize(sum(LengthKM[july < 14.73]),
            sum(LengthKM[july > 14.73]))
```

Define a cold and a warm year for the Anchor r.

```{r eval = FALSE}

# years by mean july temp

#quantiles - 7.92 and 10.16
mets %>% 
  group_by(year) %>% 
  summarize(july = mean(July_mn)) %>% 
  summarize(quantile(july, probs = c(0.1, 0.9)))

#cold 2011, warm 2004
mets %>% 
  group_by(year) %>% 
  summarize(july = mean(July_mn)) %>% 
  arrange(july)

mets %>% 
  group_by(year) %>% 
  summarize(july = mean(July_mn)) %>% 
  ggplot(aes(x = july, y = fct_reorder(factor(year), july))) +
  geom_point()

```


Thermal stress and historic metrics

```{r}
#merge historic metrics to catchments.

cat_mets_hist <- left_join(cat_spp, mets_spp %>% filter(year %in% c(2011, 2004)) %>% select(reach_id:MA7d_DAT_jd),
                           by = c("rca_id" = "reach_id")) 

cat_mets_hist_anchor <- cat_mets_hist
saveRDS(cat_mets_hist_anchor, file = "output/cat_mets_hist_anchor.rds")


```

Future metrics
```{r}
#need a long form of species and life stages. Then can filter and facet as needed.

mets_fut <- readRDS(file = "output/mets_fut.rds")

mets_fut_spp <- left_join(mets_fut, 
                          fl_spp_tbl %>% select(reach_id, LengthKM, K_s, CO_s, K_r, CO_r))


mets_fut_spp_anchor <- mets_fut_spp
saveRDS(mets_fut_spp_anchor, file = "output/mets_fut_spp_anchor.rds")

spawn_mets_fut <- mets_fut_spp %>%
  ungroup() %>% 
  select(reach_id:LengthKM, K_s, CO_s) %>% 
  pivot_longer(cols = c("K_s", "CO_s"), names_to = "Species", values_to = "present") %>% 
  filter(present > 0)

spawn_months_fut <- spawn_mets_fut %>% 
  ungroup() %>% 
  select(reach_id:Aug_mn, Species, LengthKM, scenf) %>% 
  pivot_longer(cols = June_mn:Aug_mn, names_to = "Month")

spawn_months_fut <- spawn_months_fut %>% 
  mutate(Species = factor(Species, levels = c("K_s", "CO_s"), 
                          labels = c("Chinook Salmon", "Coho Salmon")),
         Month = factor(Month, levels = c("June_mn", "July_mn", "Aug_mn"),
                        labels = c("June", "July", "August")),
         scen_cuts = cut(value, breaks = 3:18))

spawn_months_fut_anchor <- spawn_months_fut
saveRDS(spawn_months_fut_anchor, file = "output/spawn_months_fut_anchor.rds")

```




```{r}
rear_mets_fut <- mets_fut_spp %>%
  ungroup() %>% 
  select(reach_id:LengthKM, K_r, CO_r) %>% 
  pivot_longer(cols = c("K_r", "CO_r"), names_to = "Species", values_to = "present") %>% 
  filter(present > 0)

rear_months_fut <- rear_mets_fut %>% 
  ungroup() %>% 
  select(reach_id:Aug_mn, Species, LengthKM, scenf) %>% 
  pivot_longer(cols = June_mn:Aug_mn, names_to = "Month")

rear_months_fut <- rear_months_fut %>% 
  mutate(Species = factor(Species, levels = c("K_r", "CO_r"),
                          labels = c("Chinook Salmon", "Coho Salmon")),
         Month = factor(Month, levels = c("June_mn", "July_mn", "Aug_mn"),
                        labels = c("June", "July", "August")),
         scen_cuts = cut(value, breaks = 1:17)) 

rear_months_fut_anchor <- rear_months_fut
saveRDS(rear_months_fut_anchor, file = "output/rear_months_fut_anchor.rds")


```


```{r}
cat_mets_fut_anchor <- left_join(cat_spp, mets_fut, by = c("rca_id" = "reach_id"))
saveRDS(cat_mets_fut_anchor, file = "output/cat_mets_fut_anchor.rds")

```

Predictions filtered to salmon habitats only for heatmap. Need date and year and huc_12 so can create a faceted heatmap as well.

```{r}
preds <- readRDS(file = "output/preds.rds")

preds_salmon <- left_join(fl_spp_tbl %>% 
                            select(reach_id, K_s, CO_s, K_r, CO_r) %>% 
                            filter(K_s > 0 | K_r > 0 | CO_r > 0 | CO_s > 0),
                          preds)

saveRDS(preds_salmon, "output/preds_salmon.rds")

```


