---
title: "6_future predictions"
author: "Leslie Jones"
date: "1/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(rgdal)
library(zoo)
library(sf)
library(corrplot)
library(MuMIn)
library(ape)
library(readxl)
library(VIF)
library(tidyverse)
library(jtools)

library(readr)
snap <- read_csv("future/SNAP_comm_charts_export_20160926_fix_040819.csv")
View(snap)
```

## Filter on Anchor Point, AK
All 2km resolution data is downscaled using PRISM 1961-1990 baseline climatology.
Historical data is extracted from both downscaled CRU data (1961-
1990, CRU-based climatology) and also original PRISM (1961-1990) climatologies. Projected data are
based upon our 5 model average downscaled AR5 temperature (http://ckandev.
snap.uaf.edu/dataset/projected-monthly-and-derived-temperature-products-2km-cmip5-ar5) and
precipitation (http://ckan-dev.snap.uaf.edu/dataset/projected-monthly-and-derived-precipitation-products-
2km-cmip5-ar5) products, using both PRISM and CRU data as baseline climatolgies.

```{r cars}
Anchor <- snap %>% filter(community=="Anchor Point" & type=="Temperature") %>% select(community, type, daterange, scenario,resolution, junMean, julMean,  augMean, sepMean) 

#res_10 <- Anchor %>% filter(resolution=="10min") %>% mutate(jun_diff=(junMean- 9.4), jul_diff=(julMean-11.7), #aug_diff=(augMean-11.5), sep_diff=(sepMean-8.3))

res_2km <- Anchor %>% filter(resolution=="2km") %>% mutate(jun_diff=(junMean-10.0), jul_diff=(julMean-12.2), aug_diff=(augMean-11.9), sep_diff=(sepMean-8.5))

ggplot(data=res_2km) + geom_point(aes(x=daterange, y=jun_diff, colour=factor(scenario))) +theme(axis.text.x = element_text(angle = 45))

ggplot(data=res_2km) + geom_point(aes(x=daterange, y=jul_diff, colour=factor(scenario))) +theme(axis.text.x = element_text(angle = 45))

ggplot(data=res_2km) + geom_point(aes(x=daterange, y=aug_diff, colour=factor(scenario))) +theme(axis.text.x = element_text(angle = 45))

ggplot(data=res_2km) + geom_point(aes(x=daterange, y=sep_diff, colour=factor(scenario))) +theme(axis.text.x = element_text(angle = 45))
                                                                                    
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
