---
title: "Stream Temperarture Data"
author: "Leslie Jones"
date: "January 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dplyr)
library(lubridate)

###Install StreamThermal package in R

#install.packages("devtools")

library(devtools)

#install_github('tsangyp/StreamThermal')

library(StreamThermal)

```

# Wrangling Temperature Data

##Anchor River- Sue's site labeled AR-3 which is the same as CIK_14 in AKOATS and KNB.
```{r load data}
AnchorM_KNB <- read.csv(url("https://cn.dataone.org/cn/v2/resolve/urn:uuid:62f30fe8-25a3-42f6-ae23-ed0ae7a08f69", method = "libcurl"),
                    stringsAsFactors = FALSE)
```

New Anchor Rive (AR-3) data from Sue for years 2017 and 2018 to merge with above KNB.I formatted original files to column names matching KNB dataset for merge.
```{r}
AnchorM_2017 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK14_2017.csv", header=TRUE, sep=",")
AnchorM_2018 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK14_2018.csv", header=TRUE, sep=",")
AnchorM <- rbind(AnchorM_KNB, AnchorM_2017, AnchorM_2018)
AnchorM$siteID <- "CIK14"
AnchorM$AKOATS_ID <- NULL
AnchorM$UseData <- NULL
AnchorM$Freq <- 96
head(AnchorM, 10)

AnchorM$Date = AnchorM$sampleDate
AnchorM <- AnchorM%>% separate(col="Date", into=c("year", "month", "day"))
unique(AnchorM$year)
#View (AnchorM)
```

## AR-2 site from Sue

```{r}
AR2_2004 <- read.csv(file="C:/KFHP/Data/STemp/Anchor/Sue/Formatted/AR2_2004.csv", header=TRUE, sep=",")
AR2_2006 <- read.csv(file="C:/KFHP/Data/STemp/Anchor/Sue/Formatted/AR2_2006.csv", header=TRUE, sep=",")
AR2_2007 <- read.csv(file="C:/KFHP/Data/STemp/Anchor/Sue/Formatted/AR2_2007.csv", header=TRUE, sep=",")
AR2 <- rbind(AR2_2004, AR2_2006, AR2_2007)
AR2$siteID <- "AR2"
AR2$Freq <- 96
AR2$Date = AR2$sampleDate
AR2 <- AR2%>% separate(col="Date", into=c("year", "month", "day"))
unique(AR2$year)
head(AR2, 10)
```
## Two Moose site from Sue

```{r}
Moose_2004 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK1_2004.csv", header=TRUE, sep=",")
Moose_2005 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK1_2005.csv", header=TRUE, sep=",")
Moose_2006 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK1_2006.csv", header=TRUE, sep=",")
Moose_2007 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK1_2007.csv", header=TRUE, sep=",")

Moose <- rbind(Moose_2004, Moose_2005, Moose_2006, Moose_2007)
Moose$siteID <- "CIK1"
Moose$Freq <- 96
Moose$Date = Moose$sampleDate
Moose <- Moose %>% separate(col="Date", into=c("year", "month", "day"))
unique(Moose$year)
head(Moose, 10)
```
## North Fork Anchor River site from Sue 
Note that 2005 data did not have the full date range after 10/6/2005. Data was truncated. If need data post 10/6/2005 contact Sue for corrected dataset.

```{r}
NFAnchor_2004 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK2_2004.csv", header=TRUE, sep=",")
NFAnchor_2005 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK2_2005.csv", header=TRUE, sep=",")

NFAnchor <- rbind(NFAnchor_2004, NFAnchor_2005)
NFAnchor$siteID <- "CIK2"
NFAnchor$Freq <- 96
NFAnchor$Date = NFAnchor$sampleDate
NFAnchor <- NFAnchor %>% separate(col="Date", into=c("year", "month", "day"))
unique(NFAnchor$year)
head(NFAnchor, 10)
```

## Ruby Creek site from Sue

```{r}
Ruby_2004 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK3_2004.csv", header=TRUE, sep=",")
Ruby_2005 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK3_2005.csv", header=TRUE, sep=",")
Ruby_2006 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK3_2006.csv", header=TRUE, sep=",")
Ruby_2007 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK3_2007.csv", header=TRUE, sep=",")

Ruby <- rbind(Ruby_2004, Ruby_2005, Ruby_2006, Ruby_2007)
Ruby$siteID <- "CIK3"
Ruby$Freq <- 96
Ruby$Date = Ruby$sampleDate
Ruby <- Ruby %>% separate(col="Date", into=c("year", "month", "day"))
unique(Ruby$year)
head(Ruby, 10)
```

## Chakok site from Sue 
Note that 2005 data did not have the full date range after 10/6/2005. Data was truncated. If need data post 10/6/2005 contact Sue for corrected dataset.

```{r}
Chakok_2004 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK4_2004.csv", header=TRUE, sep=",")
Chakok_2005 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK4_2005.csv", header=TRUE, sep=",")
Chakok_2006 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK4_2006.csv", header=TRUE, sep=",")
Chakok_2007 <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK4_2007.csv", header=TRUE, sep=",")

Chakok <- rbind(Chakok_2004, Chakok_2005, Chakok_2006, Chakok_2007)
Chakok$siteID <- "CIK4"
Chakok$Freq <- 96
Chakok$Date = Chakok$sampleDate
Chakok <- Chakok %>% separate(col="Date", into=c("year", "month", "day"))
unique(Chakok$year)
head(Chakok, 10)
```

## Twitter Creek site from Sue

```{r}
Twitter <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK6_2005.csv", header=TRUE, sep=",")

Twitter$siteID <- "CIK6"
Twitter$Freq <- 96
Twitter$Date = Twitter$sampleDate
Twitter <- Twitter %>% separate(col="Date", into=c("year", "month", "day"))
unique(Twitter$year)
head(Twitter, 10)
```

## Beaver Creek site from Sue

```{r}
Beaver <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK7_2005.csv", header=TRUE, sep=",")

Beaver$siteID <- "CIK7"
Beaver$Freq <- 96
Beaver$Date = Beaver$sampleDate
Beaver <- Beaver %>% separate(col="Date", into=c("year", "month", "day"))
unique(Beaver$year)
head(Beaver, 10)
```

## East Creek site from Sue

```{r}
East <- read.csv(file="C:/Temperature_Database/Data/Anchor/CIK/CIK8_2005.csv", header=TRUE, sep=",")

East$siteID <- "CIK8"
East$Freq <- 96
East$Date = East$sampleDate
East <- East %>% separate(col="Date", into=c("year", "month", "day"))
unique(East$year)
head(East, 10)
```

##Merge all data into one dataset
```{r}
Anchordata <- rbind(East, Beaver, Twitter, Chakok, Ruby, NFAnchor, Moose, AR2, AnchorM)
count(Anchordata)
```

##Filter based on number of observations per day
Created a frequency variable =96 for 15 min obs. number of obs per day/freq = percentage. selected percentage >90
Then created a keep variable to be merged back to original data and filtered on original dataset.
```{r}
Anchor_filter <- Anchordata %>% group_by(siteID, Freq, sampleDate)  %>% summarize(n = n())  %>% mutate(percent = n/Freq)  %>% filter(percent > 0.90) 
head(Anchor_filter,10)
Anchor_filter$keep <-  1
Anchor2 <- Anchor_filter %>%  ungroup() %>% select ("siteID", "sampleDate", "keep")
head(Anchor2, 10)
#merge back to original datset and filter again
Anchor3 <- left_join(Anchordata, Anchor2, by = c("siteID" = "siteID", "sampleDate" = "sampleDate"))
#View(Anchor3)
Anchor_clean <- Anchor3 %>%  filter (keep==1) 

#View(Anchor_clean)
```
##Look at number of observations by month and remove all months outside of open water season June thru September
```{r}
View(Anchor_clean)
unique(Anchor_clean$year)
unique(Anchor_clean$month)
count(Anchor_clean, month, year)
Anchor_clean2 <- Anchor_clean %>% filter(month=="05"| month=="06"|month=="07" | month=="08" | month=="09")
unique(Anchor_clean2$month)
```
#Calculate daily mean, min, and max metrics
```{r}
AnchorTemp <- Anchor_clean2 %>% group_by(siteID,sampleDate, year, month, day) %>% summarize(meanT=mean(Temperature), minT=min(Temperature), maxT=max(Temperature))
View(AnchorTemp)
glimpse(AnchorTemp)
unique(AnchorTemp$year)
#write to csv
write.csv(AnchorTemp, "C:/KFHP/Data/STemp/Anchor/AnchorTemp.csv")
```


